---
title: "Filter1 Immune Cells Only Recluster"
author: "Sarah Salisbury"
output:
  html_document:
    toc: true
    number_sections: true
---

# Getting Your Data

## Setting up R Environment

### Setup
```{r setup}
# Prevent scientific notation output (e.g., we don't want very small p-values to be output as say 10^-9)
options(scipen=999)
#set seed for reproducibility
set.seed(42)
```

### Load Packages

Now load up all of the packages you'll need for this analysis

```{r load packages}
# Load up libraries
library(Seurat) # to run single cell analyses
library(ggplot2) # to make plots
library(dplyr) # to manipulate data frames
library(cowplot) # to arrange plots in a grid
library(data.table) # to use %like%
library(glmGamPoi) # helps to speed up SCTransform step
library("DoubletFinder") # to detect doublets in our dataset
```

### Load Data

```{r load rds}
AS.integrated <- readRDS("../AS.integrated_filter1.rds")


#Run PrepSCTFindMarkers, (this hasn't been done yet on the full integrated dataset and needs to be done so we can use this prior to running findmarkers on the immune subset)
AS.integrated <- PrepSCTFindMarkers(AS.integrated, assay = "SCT")
```

### Keep only good clusters

```{r remove bad clusters}
# To subset and remove clusters and keep the remaining clusters for new analysis
AS.integrated <- subset(AS.integrated, idents = c(0,4,15,19), invert = FALSE) # keep only the T-cells, myeloid, dendritic cells, neutrophils
```

### Set Default Assay

```{r set default assay}
# We're keeping the original Sctransform, so make the default the "integrated"
DefaultAssay(AS.integrated) <- "integrated"
```

## Dimensionality reduction and clustering

Now that we've integrated our data, it's time to look for cell clusters again!

We need to repeat the steps we did for each individual sample for the newly integrated dataset:
1. Generate a PCA
2. Select the best number of dimensions (PCs) to be used going forward by consulting an Elbowplot of PCs from PCA
3. Generate a UMAP
4. Cluster Cells



```{r dimensionality reduction, message = FALSE}


# 1. Linear dimension reduction
AS.integrated <- RunPCA(AS.integrated, dims = 1:10) # Please note, "dims" is not a valid argument for RunPCA, therefore the default 50 PCs were generated here in contrast to what was reported in the "Sub-Clustering" section of the Methods in the manuscript.
```
```{r dimensionality reduction visualization}
#Examine and visualize PCA results

# First print the five most informative features (nfeatures) for the first five PCA dimensions (PCs 1-5)
print(AS.integrated[["pca"]], dims = 1:5, nfeatures = 5)

# Now visualize the loadings for PCs 1 and 2 
VizDimLoadings(AS.integrated, dims = 1:2, reduction = "pca")

# Now visualize the first two axes of the PCA
DimPlot(AS.integrated, reduction = "pca", group.by = "orig.ident")

# Now visualize using a heatmap
# Note that cells and features are ordered by PCA scores.
# You can set "cells" to a value to show the most informative cells to a particular PC axis (speeds up plot generation considerably when this number is low)
# Run for the first PC now
DimHeatmap(AS.integrated, dims = 1, cells = 500, balanced = TRUE) # balanced = TRUE means you should plot an equal number of genes with both + and - scores, cells = 500 means plot the top 500 cells for this dimension

# Now repeat heatmaps for first 15 PCs
DimHeatmap(AS.integrated, dims = 1:15, cells = 500, balanced = TRUE)

# 2. Check how many PCs you need with an Elbowplot
ElbowPlot(AS.integrated, ndim = 10)
#So probably 9 PCs is ok

# 3. Make a UMAP
# use the number of dimensions as determined from Elbowplot
AS.integrated <- RunUMAP(AS.integrated, dims = 1:9, verbose = FALSE)

# Have a look at the UMAP
DimPlot(AS.integrated, reduction = "umap", group.by = "orig.ident")

# 4. Cluster cells

# First thing to do is to calculate the neighbourhood overlap (Jaccard index) between every cell and its k.param nearest neighbors (https://satijalab.org/seurat/reference/findneighbors)
# use the same number of dimensions as used for UMAP
AS.integrated <- FindNeighbors(AS.integrated, dims = 1:9)

# Now we use that neighbourhood information to figure out which cells belong together in a cluster. The resolution parameter sets the "granularity" of the clustering, so higher values will give you more clusters. This seems like a bit of a dark art. https://satijalab.org/seurat/articles/pbmc3k_tutorial.html
AS.integrated <- FindClusters(AS.integrated, resolution = 0.3) ###Potentially play with "resolution = 0.5"
```
```{r plot UMAPs, message = FALSE}
#You can save the object at this point so that it can easily be loaded back in without having to rerun the computationally intensive steps performed above, or easily shared with collaborators.
saveRDS(AS.integrated, file = "AS.integrated_filter1_immune.rds")

# Now to plot the UMAP
# note that you can set `label = TRUE` or use the LabelClusters function to help label individual clusters

# Plot UMAP with clusters in a legend
png("UMAP_clusters_filter1_immune.png", res=600, width=4200, height=3200)
DimPlot(AS.integrated , reduction = "umap", pt.size = 0.1)
dev.off()

# Plot UMAP with clusters labeled on the plot
png("UMAP_clusters_names_on_cells_filter1_immune.png", res=600, width=4200, height=3200)
DimPlot(AS.integrated, reduction = "umap", label = TRUE, pt.size = 0.1) + NoLegend()
dev.off()

# Plot UMAP with cells coloured by sample origin
png("UMAP_origin_filter1_immune.png", res=600, width=4200, height=3200)
DimPlot(AS.integrated, group.by = "orig.ident", pt.size = 0.1)
dev.off()

# Plot UMAP with clusters in a legend for each sample
png("UMAP_clusters_separated_origin_filter1_immune.png", res=600, width=6200, height=3200)
DimPlot(AS.integrated, reduction = "umap", split.by = "orig.ident", pt.size = 0.1)
dev.off()

# Plot UMAP with clusters labeled on the plot for each sample
png("UMAP_clusters_separated_origin_names_on_cells_filter1_immune.png", res=600, width=6200, height=3200)
DimPlot(AS.integrated, reduction = "umap", split.by = "orig.ident", label = TRUE, pt.size = 0.1) + NoLegend()
dev.off()

# Plot UMAP with cells coloured by cell cycle stage
png("UMAP_cellcycle_filter1_immune.png", res=600, width=4200, height=3200)
DimPlot(AS.integrated, group.by = "Phase", pt.size = 0.1)
dev.off()
```

```{r session info}
sessionInfo()
```


