---
title: "Integrating_Atlantic_and_Coho_Salmon_Data"
author: "Sarah Salisbury"
output:
  html_document:
    toc: true
    number_sections: true
---

# Getting Your Data

## Setting up R Environment

### Setup
```{r setup}
# Prevent scientific notation output (e.g., we don't want very small p-values to be output as say 10^-9)
options(scipen=999)
#set seed for reproducibility
set.seed(42) #note that UMAPs use seed of 42, not sure if setting seed here overrides this.
```

### Load Packages

Now load up all of the packages you'll need for this analysis

```{r load packages}
# Load up libraries
library(Seurat) # to run single cell analyses
library(ggplot2) # to make plots
library(dplyr) # to manipulate data frames
library(cowplot) # to arrange plots in a grid
library(data.table) # to use %like%
library(glmGamPoi) # helps to speed up SCTransform step
library("DoubletFinder") # to detect doublets in our dataset
```

## Get Data Together

### Load Data
So load up the list of Seurat objects for Atlantic and Coho Salmon
```{r load data}
# Load AS
AS.list_NoDoublets <- readRDS("AS_filter_b4_SCT/AS.list_NoDoublets.rds")

# Load CO
CO.list_NoDoublets <- readRDS("CO_filter_b4_SCT/CO.list_NoDoublets.rds")
```

### Combining AS and CO data
```{r combining AS and CO lists}
ASCO.list_NoDoublets <- c(AS.list_NoDoublets, CO.list_NoDoublets)
```

## Integration

Now that we've removed doublets we're going to integrate all of the samples together into a single dataset.
```{r Integrated feature selection, message = FALSE}
# select features that are repeatedly variable across datasets for integration
# note that nfeatures tells you how many features to include in analysis (default is 2000)
# note also that we need to specify we used SCT as our normalization method
features <- SelectIntegrationFeatures(object.list = ASCO.list_NoDoublets, nfeatures = 2000, normalization.method = "SCT")
```
```{r prepare for integration and find anchors, message = FALSE}
# Now prepare data for integration (https://satijalab.org/seurat/reference/prepsctintegration)
ASCO.list_NoDoublets <- PrepSCTIntegration(object.list = ASCO.list_NoDoublets, anchor.features = features)

# Find anchors to facilitate integration #Possibly better to use 'reduction = "rpca"' (https://satijalab.org/seurat/articles/integration_rpca.html)
anchors <- FindIntegrationAnchors(object.list = ASCO.list_NoDoublets, anchor.features = features, normalization.method = "SCT", reduction = "rpca")
```
```{r Integration, message = FALSE}
# this command creates an 'integrated' data assay
ASCO.integrated <- IntegrateData(anchorset = anchors, normalization.method = "SCT")

# specify that we will perform downstream analysis on the corrected data note that the original
# unmodified data still resides in the 'RNA' assay
DefaultAssay(ASCO.integrated) <- "integrated"
```

### Dimensionality reduction and clustering

Now that we've integrated our data, it's time to look for cell clusters again!

We need to repeat the steps we did for each individual sample for the newly integrated dataset:
1. Generate a PCA
2. Select the best number of dimensions (PCs) to be used going forward by consulting an Elbowplot of PCs from PCA
3. Generate a UMAP
4. Cluster Cells

```{r clean up memory before cell clustering again}
# Clean up memory
# At this point you may be running out of memory agai
# Ok we're going to try and remove everything now except CO.integrated to free up some memory
rm(list=setdiff(ls(), "ASCO.integrated"))
```

```{r dimensionality reduction, message = FALSE}
# 1. Linear dimension reduction
ASCO.integrated <- RunPCA(ASCO.integrated, dims = 1:50)
```
```{r dimensionality reduction visualization}
#Examine and visualize PCA results

# First print the five most informative features (nfeatures) for the first five PCA dimensions (PCs 1-5)
print(ASCO.integrated[["pca"]], dims = 1:5, nfeatures = 5)

# Now visualize the loadings for PCs 1 and 2
VizDimLoadings(ASCO.integrated, dims = 1:2, reduction = "pca")

# Now visualize the first two axes of the PCA
DimPlot(ASCO.integrated, reduction = "pca", group.by = "orig.ident")

# Now visualize using a heatmap
# Note that cells and features are ordered by PCA scores.
# You can set "cells" to a value to show the most informative cells to a particular PC axis (speeds up plot generation considerably when this number is low)
# Run for the first PC now
DimHeatmap(ASCO.integrated, dims = 1, cells = 500, balanced = TRUE) # balanced = TRUE means you should plot an equal number of genes with both + and - scores, cells = 500 means plot the top 500 cells for this dimension

# Now repeat heatmaps for first 15 PCs
DimHeatmap(ASCO.integrated, dims = 1:15, cells = 500, balanced = TRUE)

# 2. Check how many PCs you need with an Elbowplot
ElbowPlot(ASCO.integrated, ndim = 50)
#So probably 20 PCs is ok

# 3. Make a UMAP
# use the number of dimensions as determined from Elbowplot
ASCO.integrated <- RunUMAP(ASCO.integrated, dims = 1:20, verbose = FALSE)

# Have a look at the UMAP
DimPlot(ASCO.integrated, reduction = "umap", group.by = "orig.ident")

# 4. Cluster cells

# First thing to do is to calculate the neighbourhood overlap (Jaccard index) between every cell and its k.param nearest neighbors (https://satijalab.org/seurat/reference/findneighbors)
# use the same number of dimensions as used for UMAP
ASCO.integrated <- FindNeighbors(ASCO.integrated, dims = 1:20)

# Now we use that neighbourhood information to figure out which cells belong together in a cluster. The resolution parameter sets the "granularity" of the clustering, so higher values will give you more clusters. This seems like a bit of a dark art. https://satijalab.org/seurat/articles/pbmc3k_tutorial.html
ASCO.integrated <- FindClusters(ASCO.integrated, resolution = 0.2) ###Potentially play with "resolution = 0.5"
```
```{r plot UMAPs, message = FALSE}
#You can save the object at this point so that it can easily be loaded back in without having to rerun the computationally intensive steps performed above, or easily shared with collaborators.
saveRDS(ASCO.integrated, file = "ASCO.integrated.rds")

# Now to plot the UMAP
# note that you can set `label = TRUE` or use the LabelClusters function to help label individual clusters

# Plot UMAP with clusters in a legend
png("UMAP_clusters.png", res=600, width=4200, height=3200)
DimPlot(ASCO.integrated , reduction = "umap", pt.size = 0.1)
dev.off()

# Plot UMAP with clusters labeled on the plot
png("UMAP_clusters_names_on_cells.png", res=600, width=4200, height=3200)
DimPlot(ASCO.integrated, reduction = "umap", label = TRUE, pt.size = 0.1) + NoLegend()
dev.off()

# Plot UMAP with cells coloured by sample origin
png("UMAP_origin.png", res=600, width=4200, height=3200)
DimPlot(ASCO.integrated, group.by = "orig.ident", pt.size = 0.1)
dev.off()

# Plot UMAP with clusters in a legend for each sample
png("UMAP_clusters_separated_origin.png", res=600, width=6200, height=3200)
DimPlot(ASCO.integrated, reduction = "umap", split.by = "orig.ident", pt.size = 0.1)
dev.off()

# Plot UMAP with clusters labeled on the plot for each sample
png("UMAP_clusters_separated_origin_names_on_cells.png", res=600, width=6200, height=3200)
DimPlot(ASCO.integrated, reduction = "umap", split.by = "orig.ident", label = TRUE, pt.size = 0.1) + NoLegend()
dev.off()

# Plot UMAP with cells coloured by cell cycle stage
png("UMAP_cellcycle.png", res=600, width=4200, height=3200)
DimPlot(ASCO.integrated, group.by = "Phase", pt.size = 0.1)
dev.off()
```

```{r session info}
sessionInfo()
```
